---
title: "Final Project"
author: "echen642, Ben, Maggie"
date: "2/3/2022"
output: html_document
---

# Loading Packages and Data

## Loading packages

```{r Load Packages, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(ggmosaic)
library(broom)
```

## Loading Data
```{r Loading the Original Powerlifting Dataset, message=FALSE, warning = FALSE}
powerlifting <- read_csv("openpowerlifting.csv")
```

# Filtering Data

```{r Data Cleaning, message=FALSE}
# The following lines of code will filter out 885,268 athletes 
powerlifting.filtered = powerlifting %>%
  filter(Best3DeadliftKg>0,  # Filters out failed deadlift attempts
         !is.na(Best3DeadliftKg),   # Filters out athletes who didn't have any best deadlift attempts out of 3
         !is.na(Age))          # Filters out athletes with no age labeled

powerlifting.filtered = powerlifting.filtered %>%
  filter(Age > 17.5)                          # Filters our athletes less than 18 years old


# The following lines of code will filter out 333,784 athletes who are missing origin country info
powerlifting.filtered = powerlifting.filtered %>%
  filter(!is.na(Country))     # Filters out any athletes that don't have any country


## Filtering out Failed Best Squat Attempt out of 3 and Missing Best Squat Attempts
powerlifting.filtered = powerlifting.filtered %>%
  filter(Best3SquatKg >0, !is.na(Best3SquatKg))


## Filtering out Failed Best Bench Attempt out of 3 and Missing Best Bench Attempts
powerlifting.filtered = powerlifting.filtered %>%
  filter(Best3BenchKg >0, !is.na(Best3BenchKg))


## Filtering out all other countries except for USA and Russia, any athletes 25 and under, as well as adding the BigLift variable
powerlifting2.filtered = powerlifting.filtered %>%
  filter(Age > 24, Age < 41, Country %in% c("USA", "Russia")) %>%
  mutate(BigLift = Best3DeadliftKg >= mean(Best3DeadliftKg)) 

## Number of cases in RQ1 dataset
powerlifting.filtered %>%
  nrow()

## Number of cases in RQ2 dataset
powerlifting2.filtered %>%
  nrow()
```


# Graphical Representations and Numerical Summaries for RQ1

## Numerical Summaries for the Explanatory and Outcome Variables
```{r RQ1 Numerical Summaries Pt1, message=FALSE}
powerlifting.filtered %>%
  summarize(range(Age), median(Age), mean(Age), sd(Age), IQR(Age))

powerlifting.filtered %>%
  summarize(range(Best3DeadliftKg), median(Best3DeadliftKg), mean(Best3DeadliftKg), sd(Best3DeadliftKg), IQR(Best3DeadliftKg))

powerlifting.filtered %>%
  count(Equipment)

powerlifting.filtered %>%
  count(Sex)
```

## Best Deadlift Attempt out of 3 Vs Age Scatterplot's Numerical Summary
```{r RQ1 Numerical Summary Pt2, message=FALSE}
powerlifting.filtered %>%
  group_by(Equipment) %>%
  summarize(cor(Best3DeadliftKg, Age))     # Finds the correlation coefficient between "Best3DeadliftKg and "Age" among different equipment groups

powerlifting.filtered %>%  # Finds the number of missing cases for each variable in the linear regression dataset
  summarize_all(function(x) sum(is.na(x)))
```

## Best Deadlift Attempt out of 3 Vs Age Scatterplot
```{r Deadlift vs Age Scatterplot, message=FALSE}
# Using a sample of 7500 cases from the dataset that had deadlifts, age, and country filtered out, we create a scatterplot using the 2 quantitative variables "Best3DeadliftKg" and "Age", grouped by "Equipment".
powerlifting.filtered %>%
  sample_n(7500) %>%
  ggplot(aes(x = Age, y = Best3DeadliftKg, color = Equipment)) + 
  labs(x = "Age (in Years)", y = "Best Deadlift Attempt out of 3 (in Kg)") + # Labels the axis
  geom_point(alpha = 0.2) + # Alpha is a variable that controls the transparency of the points(1 = opaque and 0 = transparent)
  geom_smooth(se = FALSE, size = 1.5) + 
  scale_color_viridis_d() # Changes the colors in the scatterplot (Scale color brewer)
```

## Multiple Linear Regression Model between Best3DeadliftKg and Age
$$E[Best3DeadliftKg | Age, Equipment, Sex] = \beta_0 + \beta_1Age + \beta_2EquipmentRaw + \\ \beta_3EquipmentSinglePly + \beta_4EquipmentWraps + \beta_5SexMale$$

## 95% Confidence Interval and P-values
```{r RQ1 95% CI and P-Values, message=FALSE}
linmod = lm(Best3DeadliftKg ~ Age + Equipment + Sex, data = powerlifting.filtered)
tidy(linmod) # Let's us see the estimates of the coefficients and their p-values
confint(linmod)  # Let's us see the 95% confidence intervals of the coefficients

```

## RQ1 Model Evaluation

```{r RSquared and RSE, message=FALSE}
glance(linmod)
```

### Plotting Residuals Against Fitted Values
```{r RQ1 Residual Plots, message=FALSE}
# residuals vs fitted values (Investigating the 2 plot groups)
augment(linmod) %>%
  ggplot(aes(y = .resid, x = .fitted, color = Sex)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + # add smooth line (without gray uncertainty interval)
  geom_hline(yintercept = 0) +  # add horizontal line at y = 0
  labs(x = 'Fitted Values', y = 'Residuals', title = 'Residuals vs Fitted Values (Linear Model)') + # update axis labels
  theme_classic() # classic theme

# residuals vs fitted values
augment(linmod) %>%
  ggplot(aes(y = .resid, x = .fitted)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + # add smooth line (without gray uncertainty interval)
  geom_hline(yintercept = 0) +  # add horizontal line at y = 0
  labs(x = 'Fitted Values', y = 'Residuals', title = 'Residuals vs Fitted Values (Linear Model)') + # update axis labels
  theme_classic() # classic theme

# residuals vs explanatory (age)
augment(linmod) %>%
  ggplot(aes(y = .resid, x = Age)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + # add smooth line (without gray uncertainty interval)
  geom_hline(yintercept = 0) + # add horizontal line at y = 0
  labs(x = 'Age', y = 'Residuals', title = 'Residuals vs Age') + # update axis labels
  theme_classic()

# residuals vs explanatory (Equipment)
augment(linmod) %>%
  ggplot(aes(y = .resid, x = Equipment)) + 
  geom_boxplot() + 
  geom_smooth(se = FALSE) + # add smooth line (without gray uncertainty interval)
  geom_hline(yintercept = 0) + # add horizontal line at y = 0
  labs(x = 'Equipment', y = 'Residuals', title = 'Residuals vs Equipment') + # update axis labels
  theme_classic()

# residuals vs explanatory (Sex)
augment(linmod) %>%
  ggplot(aes(y = .resid, x = Sex)) + 
  geom_boxplot() + 
  geom_smooth(se = FALSE) + # add smooth line (without gray uncertainty interval)
  geom_hline(yintercept = 0) + # add horizontal line at y = 0
  labs(x = 'Sex', y = 'Residuals', title = 'Residuals vs Sex') + # update axis labels
  theme_classic()

# residual distribution
augment(linmod) %>% 
  ggplot(aes(x = .resid)) +
  geom_histogram()
```


# Graphical Representations and Numerical Summaries for RQ2

## Logistic Regression Model's Numerical Summaries
```{r RQ2 Numerical Summaries, message=FALSE, warning=FALSE}
# Center and spread of Age variable in the RQ2 Dataset
powerlifting2.filtered %>%
  summarize(range(Age), median(Age), mean(Age), sd(Age), IQR(Age))

# Center and spread of Age variable between Russia and US
powerlifting2.filtered %>%
  group_by(Country) %>%
  summarize(range(Age), median(Age), mean(Age), sd(Age), IQR(Age))

# Number of Cases incapable/capable of BigLift
powerlifting2.filtered %>%
  count(BigLift)

# Number of cases who Male or Female
powerlifting2.filtered %>%
  count(Sex)

# Number of cases for each Equipment Group
powerlifting2.filtered %>%
  count(Equipment)

# Number of cases from Russia and from USA
powerlifting2.filtered %>%
  count(Country)

# Number of Missing Cases for Variables in Logistic Regression Dataset
powerlifting2.filtered %>% 
  summarize_all(function(x) sum(is.na(x)))
```

## Mosaic Plot of BigLift between U.S. and Russia
```{r Mosaic Plot of BigLift between U.S and Russia, message=FALSE}
powerlifting2.filtered %>%
  ggplot() + 
  geom_mosaic(aes(x = product(BigLift, Country), fill = BigLift)) +
  facet_grid(. ~ Equipment)
```

## Logistic Regression Model
$$Log(Odds(BigLift | Country, Age, Sex, Equipment)) = \beta_0 + \beta_1CountryUSA + \beta_2Age + \beta_3SexMale + \\ \beta_2EquipmentRaw + \beta_3EquipmentSinglePly + \beta_4EquipmentWraps$$


## 95% Confidence Interval and P-Values
```{r 95% CI and P-Values, message=FALSE}
logMod = glm(BigLift ~ Country + Age + Sex + Equipment, family = binomial, data = powerlifting2.filtered)
confint(logMod) %>%
  exp()     # Let's us see the exponentiated 95% confidence interval
tidy(logMod)  # Let's us see the p-values of the coefficents
```

## Logistic Regression Model's Exponentiated Coefficients
```{r US Deadlift vs Other Countries Deadlift and Numerical Summaries, message=FALSE}
logMod %>%
  coef() %>%
  exp()  # Let's us see the exponentiated estimates of the coefficients
```

## RQ2 Model Evaluation
```{r Log Model Evaluation, message=FALSE}
threshold <- 0.25

augment(logMod, type.predict ='response') %>% ## get predicted probabilities from model
  ggplot(aes(y = .fitted, x = BigLift)) + ## compare predicted probabilites to actual outcome
  geom_boxplot() + 
  ylab("Predicted Probability of Picking a Powerlifter who can BigLift") + 
  xlab("Actual Outcome (1 = Can BigLift, 0 = Can't BigLift)") + 
  theme_classic() + 
  ggtitle('Predictions from Log Mod') +
  geom_hline(yintercept = threshold, color = "red", linetype = 2) +
  facet_wrap(.~Country)
  
  
logMod %>%
  augment(type.predict = "response") %>%
  group_by(Country, BigLift) %>%
  summarize(median(.fitted))

logMod %>%
  augment(type.predict = "response") %>%
  group_by(Country, BigLift) %>% 
  mutate(predictBigLift = .fitted >= threshold) %>% 
  count(predictBigLift, BigLift) %>% 
  mutate(correct = (BigLift == predictBigLift)) %>% 
  mutate(prop = n/sum(n)) #Specificity, False Positive Rate, False Negative Rate, Sensitivity
```
















